---
title: "Generate opportunities for win probability forecasting"
output: html_notebook
---

## Generate Dimensions
Each opportunity has following dimensions: upload date, opportunity name, customer, industry, region, marketing campaign, customer type, opportunity owner, forecast category, stage, product, price, quantity, value, estimated close date, created

```{r}
# Upload and estimated close date
created_dates <- seq.Date(from = as.Date("2017-01-01"), to = as.Date("2020-01-01"), by = "day")
# fc_dates <- seq.Date(from = as.Date("2019-10-01"), to = as.Date("2021-10-01"), by = "week")

# Opportunity name
opp_names <- c()
for (i in 1:1200){
  name <- paste(c('Opportunity', i), collapse = '_')
  opp_names[i] <- name
}

# Customer
customer_names <- c()
for (i in 1:700){
  name <- paste(c('Customer', i), collapse = '_')
  customer_names[i] <- name
}

# Industry
industry_names <- c()
for (i in 1:10){
  name <- paste(c('Industry', i), collapse = '_')
  industry_names[i] <- name
}

# Region
region_names <- c()
for (i in 1:10){
  name <- paste(c('Region', i), collapse = '_')
  region_names[i] <- name
}

# Campaign
campaign_names <- c()
for (i in 1:6){
  name <- paste(c('Campaign', i), collapse = '_')
  campaign_names[i] <- name
}

# Customer type
customer_types <- c("New Business", "Existing Business")

# Opportunity owner
owner_names <- c()
for (i in 1:15){
  name <- paste(c('Owner', i), collapse = '_')
  owner_names[i] <- name
}

# Stage
stage_names <- c("1. Marketing", "2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")

# Product
product_names <- c()
for (i in 1:12){
  name <- paste(c('Product', i), collapse = '_')
  product_names[i] <- name
}

# Fuction to generate probabilities
generate_random_probabilities <- function(n){
  nums <- runif(n, 0, 1)
  return(nums/sum(nums))
}

```


## Create the opportunities for new and existing businesses
First create new businesses, then take part of them to define as existing business. Here we add only the dimensions, which are static for a particular customer, these are: customer name, region, industry and owner.

```{r}
# Divide the opportunity names

new_business <- opp_names[1:length(customer_names)]
existing_business <- opp_names[(length(customer_names)+1):length(opp_names)]

# Create new businesses

new_business_df <- data.frame(new_business)
colnames(new_business_df) <- 'Opportunity_Name'
new_business_df$Customer <- customer_names
new_business_df$Region <- sample(x = region_names, size = length(new_business), replace = TRUE, prob = generate_random_probabilities(length(region_names)))
new_business_df$Industry <- sample(x = industry_names, size = length(new_business), replace = TRUE, prob = generate_random_probabilities(length(industry_names)))
new_business_df$Customer_Type <- 'New Business'
new_business_df$Owner <- sample(x = owner_names, size = length(new_business), replace = TRUE, prob = generate_random_probabilities(length(owner_names)))

# Create existing business
library('dplyr')
existing_business_df <- sample_n(new_business_df, length(existing_business))
existing_business_df$Customer_Type <- 'Existing Business'
existing_business_df$Opportunity_Name <- existing_business

opps_complete <- rbind(new_business_df, existing_business_df)
# print(opps_complete)

```

## Add further dimensions
Now the dimensions marketing campaign, creation date and stage for all opporunities are added
```{r}
opps_complete$Campaign <- sample(x = campaign_names, size = nrow(opps_complete), replace = TRUE, prob = generate_random_probabilities(length(campaign_names)))
stage_probabilities <- c(0.8, 0.125, 0.06, 0.005, 0.005, 0.005, 0.0, 0.0)
opps_complete$Stage <- sample(x = stage_names, size = nrow(opps_complete), replace = TRUE, prob = stage_probabilities)
opps_complete$Created <- sample(x = created_dates, size = nrow(opps_complete), replace = TRUE)

```
```{r}
library('dplyr')

opps_complete <- mutate_if(opps_complete, is.character, as.factor)
summary(opps_complete, maxsum = 20)

```
## Now create upload date which is always friday after creating
```{r}
opps_complete$DayOfWeek <- as.POSIXlt(opps_complete$Created)$wday
opps_complete$DaysToAdd <- (4 - opps_complete$DayOfWeek)%%7
opps_complete$Upload_date <- as.Date(opps_complete$Created) + opps_complete$DaysToAdd
opps_complete <- subset(opps_complete, select = -c(DayOfWeek, DaysToAdd))
print(opps_complete)

```
## Generate closing date
For lost and won opportunities "upload date" will be taken. For the 6. stage "final Bid" four weeks, for the 5. stage six weeks and so on. For the 1. stage we add 22 weeks.

```{r}
opps_complete$Close_date <- ifelse((opps_complete$Stage == "Won" | opps_complete$Stage == "Lost"), opps_complete$Upload_date, 
                                      ifelse(opps_complete$Stage == "6. Final bid", opps_complete$Upload_date + 28,
                                             ifelse(opps_complete$Stage == "5. Value Proposition", opps_complete$Upload_date + 42,
                                                    ifelse(opps_complete$Stage == "4. Identify Pains", opps_complete$Upload_date + 63,
                                                           ifelse(opps_complete$Stage == "3. Discovery", opps_complete$Upload_date + 91,
                                                                  ifelse(opps_complete$Stage == "2. Prospect", opps_complete$Upload_date + 112,
                                                                         ifelse(opps_complete$Stage == "1. Marketing", opps_complete$Upload_date + 154, NA)))))))
opps_complete$Close_date <- as.Date(opps_complete$Close_date, origin = "1970-01-01")

print(opps_complete)

```

## Add products
First, let's assign prices to products
```{r}
# Product-Price mapping
prices <-  floor(runif(length(product_names), min = 1, max = 20))
prices <- prices*100
product_prices <- data.frame(product_names, prices)
colnames(product_prices)[1] <- 'Product'
print(product_prices)
```
Now we can add products to the opportunities. To each opportunity some random number of products between 3 and 10 will be randomly chosen from the product list, with random generated probability

```{r}
# Add products
library(tidyverse)
opportunities_df <- data.frame()
for (row in 1:nrow(opps_complete)){
  line <- opps_complete[row,]
  opp_products <- sample( x = product_names, size = sample(3:10, 1), replace = FALSE, prob = generate_random_probabilities(length(product_names)))
  new_line <- data.frame(opp_products)
  colnames(new_line) <- 'Product'
  new_line$Opportunity_Name <- line$Opportunity_Name
  new_line$Customer <- line$Customer
  new_line$Region <- line$Region
  new_line$Industry <- line$Industry
  new_line$Customer_Type <- line$Customer_Type
  new_line$Owner <- line$Owner
  new_line$Campaign <- line$Campaign
  new_line$Stage <- line$Stage
  new_line$Created <- line$Created
  new_line$Upload_date <- line$Upload_date
  new_line$Close_date <- line$Close_Date
  new_line$Price <- product_prices$prices[match(new_line$Product, product_prices$Product)]
  new_line$Amount <- floor(runif(nrow(new_line), min = 1, max = 10))
  
  opportunities_df <- rbind(opportunities_df, new_line)
}


print(opportunities_df)
```

## Now let's generate the weekly updates for each opportunity. 
The data is duplicated each friday with possible changes in pipeline development. Some stages may be jumped over, e.g. after the 1. stage the opportunity may land after one week immediately in the 3. stage. The probability of that is rather low. Let's also assume, that some owner is very talented seller and sells the prodzcts faster (say, Owners 1 to 5) and one owner is rather bad (Owners 6 to 10)

```{r}
check_val <- TRUE
last_stage <- opportunities_df

# Define the probabilities for coming into next stage for worst case

sample_first_stage_wc <- c("1. Marketing", "2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_first_stage_wc <- c(0.8, 0.1, 0.03, 0.02, 0.02, 0.02, 0.002, 0.008)

sample_second_stage_wc <- c("2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_second_stage_wc <-  c(0.7, 0.1, 0.04, 0.02, 0.02, 0.02, 0.1)

sample_third_stage_wc <- c("3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_third_stage_wc <-  c(0.6, 0.12, 0.04, 0.02, 0.02, 0.2)

sample_fourth_stage_wc <- c("4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_fourth_stage_wc <-  c(0.5, 0.14, 0.04, 0.02, 0.3)

sample_fifth_stage_wc <- c("5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_fifth_stage_wc <-  c(0.4, 0.18, 0.02, 0.4)

sample_sixth_stage_wc <- c("6. Final bid", "Won", "Lost")
probability_sixth_stage_wc <-  c(0.5, 0.2, 0.3)

# Probabilities for normal case

sample_first_stage_nc <- c("1. Marketing", "2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_first_stage_nc <- c(0.6, 0.3, 0.03, 0.02, 0.02, 0.02, 0.002, 0.008)

sample_second_stage_nc <- c("2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_second_stage_nc <-  c(0.6, 0.25, 0.04, 0.02, 0.02, 0.02, 0.05)

sample_third_stage_nc <- c("3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_third_stage_nc <-  c(0.5, 0.3, 0.06, 0.02, 0.02, 0.1)

sample_fourth_stage_nc <- c("4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_fourth_stage_nc <-  c(0.5, 0.3, 0.08, 0.02, 0.1)

sample_fifth_stage_nc <- c("5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_fifth_stage_nc <-  c(0.4, 0.3, 0.2, 0.1)

sample_sixth_stage_nc <- c("6. Final bid", "Won", "Lost")
probability_sixth_stage_nc <-  c(0.3, 0.6, 0.1)

# Probabilities for best case

sample_first_stage_nc <- c("1. Marketing", "2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_first_stage_nc <- c(0.5, 0.4, 0.03, 0.02, 0.02, 0.02, 0.002, 0.008)

sample_second_stage_nc <- c("2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_second_stage_nc <-  c(0.5, 0.35, 0.04, 0.02, 0.02, 0.02, 0.05)

sample_third_stage_nc <- c("3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_third_stage_nc <-  c(0.4, 0.4, 0.06, 0.02, 0.02, 0.1)

sample_fourth_stage_nc <- c("4. Identify Pains", "5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_fourth_stage_nc <-  c(0.4, 0.4, 0.08, 0.02, 0.1)

sample_fifth_stage_nc <- c("5. Value Proposition", "6. Final bid", "Won", "Lost")
probability_fifth_stage_nc <-  c(0.3, 0.4, 0.2, 0.1)

sample_sixth_stage_nc <- c("6. Final bid", "Won", "Lost")
probability_sixth_stage_nc <-  c(0.2, 0.8, 0.0)



while(check_val){
  new_stage <- last_stage
  new_stage$Upload_date <- last_stage$Upload_date + 7
  for (opportunity in c('Opportunity_1', 'Opportunity_2')){
    opportunity_data <- new_stage[(new_stage$Opportunity_Name == opportunity),]
    # case when
    
    print(opportunity_data)
  }
  check_val = FALSE
}

# while (check_val > 0) {
# data_last_stage$Close_date <- ifelse((data_last_stage$Stage == "Won" | data_last_stage$Stage == "Lost"), data_last_stage$Upload_date, 
#                                      ifelse(data_last_stage$Stage == "7. Closing", data_last_stage$Upload_date + 14,
#                                             ifelse(data_last_stage$Stage == "6. Final bid", data_last_stage$Upload_date + 28,
#                                                    ifelse(data_last_stage$Stage == "5. Value Proposition", data_last_stage$Upload_date + 42,
#                                                           ifelse(data_last_stage$Stage == "4. Identify Pains", data_last_stage$Upload_date + 63,
#                                                                  ifelse(data_last_stage$Stage == "3. Discovery", data_last_stage$Upload_date + 91,
#                                                                         ifelse(data_last_stage$Stage == "2. Prospect", data_last_stage$Upload_date + 112,
#                                                                                ifelse(data_last_stage$Stage == "1. Marketing", data_last_stage$Upload_date + 153, NA))))))))
# data_last_stage$Close_date <- as.Date(data_last_stage$Close_date, origin = "1970-01-01")
# data_next_stage <- data_last_stage
# data_next_stage$Upload_date <- data_next_stage$Upload_date + 7
# #rint(data_next_stage)
# 
# first_stage <- subset(data_next_stage, data_next_stage$Stage == "1. Marketing")
# second_stage <- subset(data_next_stage, data_next_stage$Stage == "2. Prospect")
# third_stage <- subset(data_next_stage, data_next_stage$Stage == "3. Discovery")
# fourth_stage <- subset(data_next_stage, data_next_stage$Stage == "4. Identify Pains")
# fifth_stage <- subset(data_next_stage, data_next_stage$Stage == "5. Value Proposition")
# sixth_stage <- subset(data_next_stage, data_next_stage$Stage == "6. Final bid")
# seventh_stage <- subset(data_next_stage, data_next_stage$Stage == "7. Closing")
# closed_stage <- subset(data_next_stage, (data_next_stage$Stage == "Won") | (data_next_stage$Stage == "Lost"))
# 
# first_stage_austria <- first_stage[(first_stage$Region == "Austria"),]
# first_stage_rest <- first_stage[(first_stage$Region != "Austria"),]
# first_stage_austria$Stage <- sample(x = c("1. Marketing", "2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "7. Closing", "Won", "Lost"), size = dim(first_stage_austria)[1], replace = TRUE, prob = c(0.8, 0.1, 0.02, 0.02, 0.02, 0.02, 0.01, 0.002, 0.008))
# first_stage_rest$Stage <- sample(x = c("1. Marketing", "2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "7. Closing", "Won", "Lost"), size = dim(first_stage_rest)[1], replace = TRUE, prob = c(0.63, 0.2, 0.08, 0.02, 0.02, 0.02, 0.02, 0.005, 0.005))
# 
# second_stage_austria <- second_stage[(second_stage$Region == "Austria"),]
# second_stage_rest <- second_stage[(second_stage$Region != "Austria"),]
# second_stage_austria$Stage <- sample(x = c("2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "7. Closing", "Won", "Lost"), size = dim(second_stage_austria)[1], replace = TRUE, prob = c(0.8, 0.09, 0.02, 0.01, 0.01, 0.01, 0.01, 0.05))
# second_stage_rest$Stage <- sample(x = c("2. Prospect", "3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "7. Closing", "Won", "Lost"), size = dim(second_stage_rest)[1], replace = TRUE, prob = c(0.5, 0.3, 0.1, 0.02, 0.02, 0.02, 0.02, 0.02))
# 
# third_stage_austria <- third_stage[(third_stage$Region == "Austria"),]
# third_stage_rest <- third_stage[(third_stage$Region != "Austria"),]
# third_stage_austria$Stage <- sample(x = c("3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "7. Closing", "Won", "Lost"), size = dim(third_stage_austria)[1], replace = TRUE, prob = c(0.8, 0.05, 0.05, 0.05, 0.01, 0.01, 0.03))
# third_stage_rest$Stage <- sample(x = c("3. Discovery", "4. Identify Pains", "5. Value Proposition", "6. Final bid", "7. Closing", "Won", "Lost"), size = dim(third_stage_rest)[1], replace = TRUE, prob = c(0.4, 0.4, 0.08, 0.04, 0.02, 0.02, 0.02))
# 
# fourth_stage_austria <- fourth_stage[(fourth_stage$Region == "Austria"),]
# fourth_stage_rest <- fourth_stage[(fourth_stage$Region != "Austria"),]
# fourth_stage_austria$Stage <- sample(x = c("4. Identify Pains", "5. Value Proposition", "6. Final bid", "7. Closing", "Won", "Lost"), size = dim(fourth_stage_austria)[1], replace = TRUE, prob = c(0.7, 0.15, 0.06, 0.02, 0.02, 0.05))
# fourth_stage_rest$Stage <- sample(x = c("4. Identify Pains", "5. Value Proposition", "6. Final bid", "7. Closing", "Won", "Lost"), size = dim(fourth_stage_rest)[1], replace = TRUE, prob = c(0.3, 0.5, 0.1, 0.06, 0.02, 0.02))
# 
# fifth_stage_austria <- fifth_stage[(fifth_stage$Region == "Austria"),]
# fifth_stage_rest <- fifth_stage[(fifth_stage$Region != "Austria"),]
# fifth_stage_austria$Stage <- sample(x = c("5. Value Proposition", "6. Final bid", "7. Closing", "Won", "Lost"), size = dim(fifth_stage_austria)[1], replace = TRUE, prob = c(0.7, 0.2, 0.06, 0.02, 0.02))
# fifth_stage_rest$Stage <- sample(x = c("5. Value Proposition", "6. Final bid", "7. Closing", "Won", "Lost"), size = dim(fifth_stage_rest)[1], replace = TRUE, prob = c(0.3, 0.5, 0.1, 0.08, 0.02))
# 
# sixth_stage_austria <- sixth_stage[(sixth_stage$Region == "Austria"),]
# sixth_stage_rest <- sixth_stage[(sixth_stage$Region != "Austria"),]
# sixth_stage_austria$Stage <- sample(x = c("6. Final bid", "7. Closing", "Won", "Lost"), size = dim(sixth_stage_austria)[1], replace = TRUE, prob = c(0.6, 0.1, 0.1, 0.2))
# sixth_stage_rest$Stage <- sample(x = c("6. Final bid", "7. Closing", "Won", "Lost"), size = dim(sixth_stage_rest)[1], replace = TRUE, prob = c(0.4, 0.3, 0.2, 0.1))
# 
# seventh_stage_austria <- seventh_stage[(seventh_stage$Region == "Austria"),]
# seventh_stage_rest <- seventh_stage[(seventh_stage$Region != "Austria"),]
# seventh_stage_austria$Stage <- sample(x = c("7. Closing", "Won", "Lost"), size = dim(seventh_stage_austria)[1], replace = TRUE, prob = c(0.4, 0.3, 0.3))
# seventh_stage_rest$Stage <- sample(x = c("7. Closing", "Won", "Lost"), size = dim(seventh_stage_rest)[1], replace = TRUE, prob = c(0.2, 0.6, 0.2))
# 
# 
# data_next_week <- do.call("rbind", list(first_stage_austria, first_stage_rest, second_stage_austria, second_stage_rest, third_stage_austria, third_stage_rest, fourth_stage_austria, fourth_stage_rest, fifth_stage_austria, fifth_stage_rest, sixth_stage_austria, sixth_stage_rest, seventh_stage_austria, seventh_stage_rest, closed_stage))
# 
# data_next_week$Stage <- ifelse((data_next_week$Region == "Austria" & data_next_week$Product == "MODL_BI_FORC_SAAS" & data_next_week$Stage != "1. Marketing"), "1. Marketing", data_next_week$Stage)
# data_next_week$Stage <- ifelse((data_next_week$Region == "Austria" & data_next_week$Product == "MODL_BI_FORC_SAAS" & data_next_week$Stage != "2. Prospect"), "2. Prospect", data_next_week$Stage)
# 
# 
# test_data <- rbind(test_data, data_next_week)
# data_last_stage <- data_next_week
# check_data <- data_next_week[(data_next_week$Upload_date < "2019-08-30"),]
# check_val <- dim(check_data)[1]
# print(check_val)
# 
# }
# print(test_data)

```


